---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# toyrcpp2

<!-- badges: start -->
<!-- badges: end -->

The goal of `toyrcpp2` is to demonstrate how to create an R package that uses `Rcpp`. The functions in this package are not particularly useful in and of themselves. See the package vignettes for documentation that walks through how to create a package with Rcpp.

## Installation

You can install the development version of `toyrcpp2` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("GrindeLab/toyrcpp2", build_vignettes = TRUE)
```

## Example

This is a basic example which shows you how to use the R and C++ functions contained within this package:

```{r example}
library(toyrcpp2)
timesTwoR(455)
timesTwoCpp(455)
```

The two functions should give you the exact same result since they implement the same task; one is just written using R code (`timesTwoR`) and the other uses C++ code (`timesTwoCpp`) which is called within R thanks to the `Rcpp` package.

As mentioned above, the most useful aspect of this package is the vignette. Take a look at the `creating-rcpp-package` vignette for a description of the steps you need to take to create an R package using Rcpp.

```{r, eval = FALSE}
browseVignettes("toyrcpp2")
```
